---
title: "Lab 2"
author: "kaori hirano"
date: "6 10 2023"
subtitle: "Due: 11:59 PM on 6/15/2023"
format: pdf
---
```{r packages}
library(tidyverse)
library(lubridate)
```
# Part 1

## Q1

```{r dataframe construction}

date <- mdy(c(071220, 71320, 71420, 71520, 71620,71720,71820,71920, 72020,72120,72220,72320,72420,72520,72620,72720,72820, 72920,73020,73120,80120,80220,80320))

mask <- (c(25.5, 19.5, 19.5, 20.5, 19.5, 19.5, 20.5, 20, 20.5, 21.4, 19.5, 19.5, 20.5, 19, 19.4, 17, 16.5, 16.7, 16.7, 16.5, 16.6, 16.5, 16.5))

noMask <- (c(9.9, 9.1, 9.3, 9.9, 9.9, 9.6, 9.6, 9, 8.5, 8.6, 8.5, 9.8, 9.9, 10, 9.8, 9.8, 9.8, 10, 9.6, 9, 9.1, 9, 9.2))

(cbind(length(noMask), length(mask), length(date)))
d <- data.frame(time(date), mask, noMask)

```


## Q2 

```{r making graph}
ggplot(d, 
       aes(x = date, y = caseNumber, color = maskStatus, group = maskStatus)) +
  geom_line() +
  facet_wrap(~ subgroup)

ggplot(d, aes(x = date)) + 
  geom_line(aes(y = mask, color = 'Mask Mandate')) + 
  geom_line(aes(y = noMask, color = 'No Mask Mandate')) +
  scale_color_manual(values = c("lightsalmon1", "aquamarine3")) + 
  labs( 
    x = "Date", y = "Daily Cases/Per 100K Population", 
    color = NULL, 
    title = "Kansas COVID 7 Day Rolling Average of Daily Cases/Per 100K Population by Mask vs No Mask Mandate Counties", 
    caption = "Source: Kansas Department of Health and Environment" 
  ) 
```

## Q3
In this visual it is more clear that there is a higher incidence of covid cases per 100K people in mask mandate counties than non mask mandate counties from July 11-August 3 2020 in Kansas. 

## Q4
It doesn't necessarily give us useful information regarding mask wearing and COVID that can be used based on this alone or immediantely, but it gives evidence that mask mandates may not be as effective at preventing the spread of covid as expected given the lower incidence rates of non-mask mandated counties. It invites further research. One thing I would be interested to see included is the incidence rate by county when population and/or population density is included. 



...

# Part 2

## Q1
```{r p2q1}
p <- read.csv('data/paygap.csv') |>
  janitor::clean_names()        # just for practice
typeof(p$due_date)
p <- p |> mutate(year = year(as_datetime(due_date)) - 1)
p$year
```

## Q2
```{r p2q2}
count(p, vars = year) #, wt_var = NULL) # gets reports per year

length(p$employer_id) - length(unique(p$employer_id)) # subtracts all employers
# from duplicates removed employers list length
```
There are far more observations in 2017-2021 than in 2022. 2019 also has fewer, 
by only by around 3000 rather than 9000 responses. This could be because of the year
of data we are using. After looking at the codebook, it shows that the most
recent one is 2023-2024. This suggests that the employers listed in 2022 may have 
reported earlier than the deadline so their numbers for the upcoming year are already listed. 

There are 35963 unique companies reporting in the dataset. 

## Q3
```{r p2q3 visualizing pay gap}
# Is there any evidence that the pay gap decreased over the time period covered by this
# data set? Use at least one visualization to answer this question
plot <- plot(p$year, p$diff_median_hourly_percent, 
             xlab = 'Year',
             ylab = 'Median Hourly Difference In Pay (Percent)',
             main = 'Gender Pay Gap in the UK by Year')
```
Based on the visualization, there does appear to be a (descriptive) decrease in 
the overall range of the gap, with more data being clustered toward more equal
pay (toward 0) in later years, especially 2022. 

## Q4
```{r p2q4}
levels(as.factor(p$employer_size))
p <- p |> 
  mutate(employer_size = 
           fct_relevel(employer_size, c('Less than 250', '250 to 499', 
                                        '500 to 999', '1000 to 4999', 
                                        '5000 to 19,999', '20,000 or more', 
                                        'Not Provided'))) 
# recoding so in order of inc size
       
(plot1 <- ggplot(p, aes(x = employer_size, y = diff_median_hourly_percent)) + 
  geom_boxplot(fill = 'salmon') +
   labs( 
    x = "Employer Size", 
    y = 'Median Hourly Difference In Pay (Percent)',
    title = 'Gender Pay Gap in the UK by Employer Size'))

(plot2 <- ggplot(p, aes(x = employer_size, y = diff_median_hourly_percent)) +  #plot without outliers
  geom_boxplot(fill = 'salmon', outlier.shape = NA) +
  coord_cartesian(ylim=c(quantile(p$diff_median_hourly_percent, prob = .25) - 50, quantile(p$diff_median_hourly_percent, prob = .80) + 50)) + 
   labs( 
    x = "Employer Size", 
    y = 'Median Hourly Difference In Pay (Percent)',
    title = 'Gender Pay Gap in the UK by Employer Size (no outliers)'))
```
Descriptively speaking, there appears to be slightly less of a pay gap in employers
with a size of 20,000 or more, as seen by the median which is slightly closer to
zero than the other employer sizes, by around 5%. 


























